---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit README.Rmd. -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "tools/README-"
)
```

<!-- badges: start -->
[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/)
[![Contributor Covenant](https://img.shields.io/badge/Contributor%20Covenant-2.1-4baaaa.svg)](code_of_conduct.md) 
<!-- badges: end -->


# theHUB: the Hub Utility Box.

_**the Hub Utility Box**_ (_**theHUB**_), a collection of analysis tools composed of functions, scripts, electronic notebooks, and worked examples. _theHUB_ is how we will share these tools with the MSU community to improve the analyses of student outcomes, student success, and survey responses. A standard set of processes and methods along with a defined lexicon enables reproducibility, reduces the time needed to complete analysis-based projects, and allows the quick addition of new methods while encouraging data exploration. Our desire is for theHUB’s functionality to grow over time, and for others at MSU to contribute new features and methods for the MSU community. Developed and implemented in R -- an open-source application with a robust community (on campus and around the world) -- _theHUB_ removes the need for a software license while providing access to a myriad of statistical analysis and graphing options. _theHUB’s_ documentation allows anyone to use and understand the tools. Initially, _theHUB_ will contain three Learning Analytics v2.0 projects: the survey analysis package presented at the 2021 Spring Conference, _Survey Connections_, along with the _in Semester Analysis Aiding/Assisting Students_, and _Unified Student Success and Outcomes Analysis Reports_ packages; each package is briefly described below. We expect _theHUB_ to grow as we add new, yet to be conceived, ideas and analysis methods from the literature, our conversations with others around campus, and conferences and seminars. 

_**Survey Connections**_: _Analyze survey responses to gain a better understanding of the relationships between responses and respondents._ Traditionally, survey analysis returns a collection of response counts and comment summaries without considering important and insightful relationships between questions and responses. The _Survey Connections_ package provides workflows for cleaning survey data and performing Likert and Likert-like, multi-response, and comment (text) analyses. The package’s goal is to easily allow the discovery of relationships between responses and the creation of cohorts for further exploration and analysis.

_**in Semester Analysis Aiding/Assisting Students**_ (_**iSAAS**_): _Using a student’s in-semester performance and course participation data provides educators a method to identify those needing academic support._ Identifying the area(s) within a course that a student needs assistance during the semester significantly improves the student’s chance of a favorable outcome. Currently, educators identify the area(s) of needed improvement via MSU’s Enhancing Academic Success Early (EASE) reports. Automating the identification of students during the semester using gradebook data allows educators and academic units to provide students with the required resources to improve their course standing, thus leading to better student outcomes and success. However, we understand that constructing detailed EASE reports can be time-consuming, especially for educators asked to do more with fewer resources. _iSAAS_ builds on the ideas and goals of the current EASE reporting system and is an automated system to construct detailed reports using D2L course gradebook data, reducing educator load. In addition, the automated nature of iSAAS allows educators to provide earlier and more frequent reports to the student. Initially, _iSAAS_ will focus on large-enrollment, first and second-year gateway courses.

_**Unified Student Success and Outcomes Analysis Reports**_ (_**USOAR**_): _Provide academic units with unit-level learning analytics reports while allowing them to explore student success analysis specific to their needs and interests._ Academic units (College or Department) have a collection of student-centric analyses they perform to understand their students, the student’s success and outcomes, and evaluate the unit’s academic endeavours. One expects overlaps between the units’ analyses, yet the breadth and depth of the analysis likely differ, creating an analytics gap. The analytics gap turns into an analytics haves and have-nots situation because few units on campus have full-time analytics groups. The analysis-chasm between units will continue to expand without a resource to provide student success and outcomes analysis. A centralized resource will also reduce the significant logistical burden for units by leveraging their collective knowledge into a common analytics framework that benefits all academic units. A unified learning analytics platform allows new analytics methods to quickly propagate to the MSU community while reducing the analytics gap for those without dedicated student success and outcomes analysis groups. The _USOAR_ system is based on the belief in an egalitarian analytics ecosystem where everyone (students and academic units) benefits.



## Installing `theHUB`

`theHUB` is _only_ available on [GitHub](https://github.com/emilioxavier/theHUB/). To install it:

```{r, eval=FALSE}
# Install from GitHub:
install.packages("devtools")
devtools::install_github("emilioxavier/theHUB")
```


## How to use `theHUB`

```{r}
library(theHUB)
```


## Have a suggestion? Need help? Found a bug?

* Contact [Emilio](https://github.com/emilioxavier/) at emilio@msu.edu OR emilio.esposito@gmail.com
* Submit an [issue via GitHub](https://github.com/emilioxavier/theHUB/issues/)


## Code of conduct

Please note that this project is released with a [Contributor Code of Conduct](https://github.com/emilioxavier/theHUB/blob/master/CONDUCT.md). By participating in this project you agree to abide by its terms.

## License

MIT © [Emilio Xavier Esposito](https://github.com/emilioxavier/)
