---
title: "Dataset: Country and Currency Dataset"
date: "Updated 03/Mar/2023"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Dataset: Country and Currency Dataset}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo=FALSE, eval=FALSE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(knitr)
library(kableExtra)
```


## Country and Currency Dataset

The Country and Currency Dataset was constructed to help the [Predictive Analytics Group](https://thehopefulbox.com/articles/web_only/PAG_About.html) (PAG) explore the impact of currency exchange rates on applicants. The [Office of Admissions](https://admissions.msu.edu/) at [Michigan State University](https://msu.edu) (MSU) uses [slate](https://technolutions.com/solutions/slate) for applicants management. Unfortunately, the country names within slate do not match the [ISO 3166](https://en.wikipedia.org/wiki/ISO_3166) standard names. (_NB_: This is a standard and common problem across a multitude of organizations and applications.) To remedy this inconsistency, the following process was performed; code provided below.

 - Determine the unique countries within the slate data.
 - Extract the ISO 3166 information from `ISOcodes::ISO_3166_1`.
 - Extract the currency name and code from `priceR::currencies()`.
   - Extract the country two-letter code from the currency code.
 - Merge the ISO 3166 information and the currency information using the two-letter code.
 - Merge the `ISOcodes::ISO_3166_1` countries with the slate country using the **Name** column.
 - Export merged information to an Excel workbook.
 - Within Excel, reorder data based on the **Name** column.
   - Instances where the slate name does not match the ISO 3166 name are easily recognizable because there are no **Alpha_2**, **Alpha_3**, **Numeric**, and **Name** entries, move the two slate cells (**Name.SLATE** and **SLATE.tf**) to the appropriate ISO 3166 row.
   - Wikipedia was used to determine currency used in the countries without a noted currency. See [Data Sources](#DataSources).
 - Import the updated Excel workbook as a tibble.



## Construction of the Country and Currency Dataset
The following are the commands used to construct the Country and Currency Dataset broken into individual steps. The editing within the Excel workbook is not included. 
```{r loadPackages, eval=FALSE, include=TRUE}
library(theHUB)
library(ISOcodes)
library(priceR)
```


### Extract Country Names Information
```{r ExtractCountry, eval=FALSE, include=TRUE}
country.iso <- as_tibble(ISOcodes::ISO_3166_1) |>
  mutate(Name.ISO=Name)
```

### Extract Country and Currency Information
```{r ExtractCurrency, eval=FALSE, include=TRUE}
country.currencies <- priceR::currencies() |>
  as_tibble() |>
  separate(col="code", into=c("Alpha_2", "currency.DROP"), remove=FALSE, sep="[[:alpha:]]{1}$") |>
  select(-currency.DROP) |>
  rename("currency"="description",
         "currency.code"="code")
```

### Merge All the Information 
```{r MergeEverything, eval=FALSE, include=TRUE}
country.DATA <- full_join(x=country.iso, y=country.currencies, by="Alpha_2") |>
  full_join(y=SLATE.country.names, by="Name")
```


### Write to Excel workbook
```{r WriteXLSX, eval=FALSE, include=TRUE}
WriteXLS::WriteXLS(x="country.DATA",
                   ExcelFileName="CountryISO-slate_matches.xlsx",
                   FreezeRow=1)
```


## The Dataset

The 257 International Organization for Standardization (ISO) recognized countries and the currencies they use.


```{r loadPackages2, include=FALSE}
library(tidyselect)
library(dplyr)
library(tidyr)
```


```{r theDataset, echo=FALSE}
country.currency.DISPLAY <- select(theHUB::country.currency, -Common_name, -contains("SLATE")) |>
  mutate(across(where(is.character), ~replace_na(.x, "")))
col.names <- c("Alpha 2", "Alpha 3", "Numeric", "Name", "Official Name", "Name (ISO)", "Currency", "Currency Code")

kableExtra::kbl(country.currency.DISPLAY, booktabs=TRUE, caption=NULL, col.names=col.names) |> 
  kableExtra::kable_styling(latex_options="striped")
```


## Data Sources {#DataSources}
 - List of countries (via `ISOcodes::ISO_3166_1`) from ISOcodes ([CRAN](https://cran.r-project.org/web/packages/ISOcodes/))
 - List of currencies (via `priceR::currencies()`) from priceR ([CRAN](https://cran.r-project.org/package=priceR) & [GitHub](https://github.com/stevecondylios/priceR))
 - Wikipedia's [List of circulating currencies](https://en.wikipedia.org/wiki/List_of_circulating_currencies)
 - Individual currency [Wikipedia](https://en.wikipedia.org/) pages (too many to list)
 - Individual country [Wikipedia](https://en.wikipedia.org/) pages (also, too many to list)


