---
title: "How To: Creating Donut Plots"
date: "Updated 20/Dec/2022"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How To: Creating Donut Plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, echo=FALSE, eval=FALSE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(knitr)
```

Donut[^1] plots (also known as ring plots) are an efficient way to display data that represents proportions of things within a dataset.

[^1]: You can also spell donut as doughnut. Those unfamiliar with these tasty concoctions, see the [Wikipedia entry](https://en.wikipedia.org/wiki/Doughnut) or visit your nearest [donut shop](). If you live in East Lansing, I highly recommend [Groovy Donuts](https://www.groovydonuts.com/)!


## Constructing the Donut Plot Data

Using the [mpg dataset](https://ggplot2.tidyverse.org/reference/mpg.html) from [ggplot2](https://ggplot2.tidyverse.org/), we will explore the construction of the donut `data.frame`. 


The `class` column will be used for facets.

```{r constructDataFrame, eval=FALSE, include=FALSE}

library(tidyverse)
library(theHUB)


data <- mpg; col.oi="cyl"; facetBy="class"; layerBy=NULL; category.order="count"; col.count=NULL; r.inner=4; r.outer=6

make.donut.data(data=mpg, col.oi="cyl", facetBy="class")


```



## Donut Plot



## A Donut Plot with Layers



## Faceted Donut Plots



## Faceted Donut Plots with Layers

