---
title: "How To: Creating Donut Plots"
date: "Updated 20/Dec/2022"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How To: Creating Donut Plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, echo=FALSE, eval=FALSE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(knitr)
```

Donut[^1] plots (also known as ring plots) are an efficient way to display data that represents proportions of things within a dataset.

[^1]: You can also spell donut as doughnut. Those unfamiliar with these tasty concoctions, see the [Wikipedia entry](https://en.wikipedia.org/wiki/Doughnut) or visit your nearest [donut shop](). If you live in East Lansing, I highly recommend [Groovy Donuts](https://www.groovydonuts.com/)!


## Constructing the Donut Plot Data

Using the [mpg dataset](https://ggplot2.tidyverse.org/reference/mpg.html) from [ggplot2](https://ggplot2.tidyverse.org/), we will explore the construction of the donut `data.frame`. 


The `class` column will be used for facets.

```{r constructDataFrame, eval=FALSE, include=FALSE}

library(tidyverse)
library(theHUB)

data=mpg; category="cyl"; facetBy=NULL; category.order="count"; category.count=NULL; r.inner=4; r.outer=6; levels.rev=FALSE; layerBy=NULL; layer.order <- c("4", "r", "f")

data=mpg; category="cyl"; facetBy="class"; category.order="count"; category.count=NULL; r.inner=4; r.outer=6; levels.rev=FALSE; layerBy="drv"; layer.order <- c("4", "r", "f")
data=mpg; category="cyl"; facetBy="class"; category.order="count"; category.count=NULL; r.inner=4; r.outer=6; levels.rev=FALSE; layerBy=NULL; layer.order <- c("4", "r", "f")

make.donut.data(data=mpg, category="cyl", facetBy="class", layerBy="drv")
make.donut.plot(donut.DATA=donut.DATA, colour.palette=msu.palette, label.col="Labels.short", facet.ncol=3)
make.donut.plot(donut.DATA=donut.DATA, colour.palette=msu.palette, label.col="Labels.short")


```



## Custom Colour Palette
When creating a collection of donut plots in a facet it is important to have the same colours assigned to the same segments for all donut plots. Often, each donut plot does not always include the same collection of segments. The following code ensures that the same segment-colour combination is used for each donut plot.

```
custom.palette <- msu.palette[1:9]
names(custom.palette) <- all.donut.DATA$Categories |> unique()
```

If a "named colour palette" is not provided, one is created from the categories within the `donut.DATA` data.frame and the provided `colour.palette` to ensure a consistent colour palette is used for all donut plots.




## Donut Plot
library(tidyverse)
library(theHUB)

data=mpg; category="cyl"; facetBy=NULL; category.order="count"; category.count=NULL; r.inner=4; r.outer=6; levels.rev=FALSE; layerBy=NULL; layer.order <- c("4", "r", "f")

## !!! works !!!!
donut.DATA <- make.donut.data(data=mpg, category="cyl", facetBy=NULL, layerBy=NULL, category.order="count", category.count=NULL, r.inner=4, r.outer=6, levels.rev=FALSE)
make.donut.plot(donut.DATA=donut.DATA, colour.palette=msu.palette, label.col="Labels.short", facet.ncol=3)

## !!! works !!!!
donut.DATA <- make.donut.data(data=mpg, category="cyl", facetBy="class", layerBy=NULL, category.order="count", category.count=NULL, r.inner=4, r.outer=6, levels.rev=FALSE)
make.donut.plot(donut.DATA=donut.DATA, colour.palette=msu.palette, label.col="Labels.short", facet.ncol=3)

## !!! works !!!!
donut.DATA <- make.donut.data(data=mpg, category="cyl", facetBy="class", layerBy="drv", category.order="category", layer.order="descend", category.count=NULL, r.inner=4, r.outer=6, levels.rev=FALSE)
make.donut.plot(donut.DATA=donut.DATA, colour.palette=msu.palette, label.col="Labels.short", facet.ncol=3)

## !!! works !!!!
donut.DATA <- make.donut.data(data=mpg, category="cyl", facetBy=NULL, layerBy="drv", category.order="category", layer.order="descend", category.count=NULL, r.inner=4, r.outer=6, levels.rev=FALSE)
make.donut.plot(donut.DATA=donut.DATA, colour.palette=msu.palette, label.col="Labels.short", facet.ncol=3)


## A Donut Plot with Layers
data=mpg; category="cyl"; facetBy=NULL; layerBy="drv"; category.order="count"; layer.order="descend"; category.count=NULL; r.inner=4; r.outer=6; levels.rev=FALSE



## Faceted Donut Plots



## Faceted Donut Plots with Layers

